---
title: 'Tell the difference: Pop or Country, Hiphop or R&B?'
author: "Huizhe (Sunny) ZHU"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

Not many people can tell the difference between Country music and Pop music. Pop means 'popular', while country music is also popular in many regions including the US. Meanwhile, Rhythm & Blues (R&B) is considered as a mix of Pop, Hip-Hop etc. It is difficult to distinguish Pop or Country, Hiphop or R&B. 

Given the database that contains more than 120,000 songs, it is exciting to identify the differences between these genres. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Step 0 - Load libraries
Install packages & Load libraries
```{r load libraries, message=FALSE, warning=FALSE, include = FALSE}
install.packages('broom')
library(tm)
library(data.table)
library(tidytext)
library(tidyverse)
library(DT)
```

#### Step 1 - Load the data
```{r, include = FALSE}
# load lyrics data
library(readr)
raw_lyrics <- read.csv("lyrics.csv", stringsAsFactors = FALSE)

```


#### Step 2 - Preliminary cleaning of text

Convert letters to the lower case, remove punctuation, numbers, empty words and extra white space.
```{r text processing in tm}
# function for removimg leading and trailing whitespace from character strings 
leadingWhitespace <- content_transformer(function(x) str_trim(x, side = "both"))

# remove stop words
data("stop_words")
word <- c("lot", "today", "months", "month", "wanna", "wouldnt", "wasnt", "ha", "na", "ooh", "da",
        "gonna", "im", "dont", "aint", "wont", "yeah", "la", "oi", "nigga", "fuck",
          "hey", "year", "years", "last", "past", "feel")
stop_words <- c(stop_words$word, word)

```


```{r}
# clean the data and make a corpus
corpus <- VCorpus(VectorSource(raw_lyrics$lyrics))%>%
  tm_map(content_transformer(tolower))%>%
  tm_map(removePunctuation)%>%
  tm_map(removeWords, character(0))%>%
  tm_map(removeWords, stop_words)%>%
  tm_map(removeNumbers)%>%
  tm_map(stripWhitespace)%>%
  tm_map(leadingWhitespace)
```


#### Step 3 - Stemming and converting to tidy object

Stemming reduces a word to its word *stem*. Each row contains 1 song - its stemmed words
* Note: 11 empty lines 
```{r stemming}
stemmed <- tm_map(corpus, stemDocument) %>%
  tidy() %>%
  select(text)
```


#### Step 4 - Tokenization on original words

*9.51* million rows (words) in total, one column
```{r tidy dictionary}
dict <- tidy(corpus) %>%
  select(text) %>%
  unnest_tokens(dictionary, text)
```


#### Step 5 - Tokenization on stemmed words

Add ID = row_number, dict = original word

*9.51* million rows, 3 columns
```{r tidy stems with dictionary}
completed <- stemmed %>%
  mutate(id = row_number()) %>%
  unnest_tokens(stems, text) %>%
  bind_cols(dict) 
```


#### Step 6 - Stem completion

Picking the original word (from the same root) with the highest frequency. The new column created -  'word' - will be used to replace stem, which is sometimes hard for people to understand. 
```{r stem completion, warning=FALSE, message=FALSE}
completed1 <- completed %>%
  group_by(stems) %>%
  count(dictionary) %>%
  mutate(word = dictionary[which.max(n)]) %>%
  ungroup() %>%
  select(stems, word) %>%
  distinct() %>%
  right_join(completed) %>%
  select(-stems)
```


#### Step 7 - Pasting stem completed individual words into their respective lyrics

Put 'word' generated from Step 6, back to each song row - each row contains words like before - but the words are transforme, from -> stemmed -> completed stem
```{r reverse unnest}
completed2 <- completed1 %>%
  group_by(id) %>%
  summarise(stemmedwords= str_c(word, collapse = " ")) %>%
  ungroup()
```


#### Step 8 - Keeping a track of the processed lyrics with their own ID
```{r cleaned hm_data, warning=FALSE, message=FALSE}
raw_lyrics1 <- raw_lyrics %>%
  mutate(id = row_number()) %>%
  inner_join(completed2)
```


#### Step 9 - Exporting the processed text data into a CSV file
```{r export data}
save(raw_lyrics1, file="../output/processed_lyrics3.RData")
```

The final processed data is ready to be used for any kind of analysis.






# Songs Classification - Country vs Pop, Hiphop vs R&B

#### Step 1 - Create sub-dataset
In the dataset, there are 7534 Country songs, 18697 Pop songs,  8905 HipHops, and 2174 R&B. 
```{r}
table(raw_lyrics1$genre)
```


```{r}
country = raw_lyrics1[raw_lyrics1$genre == 'Country',]
pop = raw_lyrics1[raw_lyrics1$genre == 'Pop',]
hiphop = raw_lyrics1[raw_lyrics1$genre == 'Hip-Hop',]
RB = raw_lyrics1[raw_lyrics1$genre == 'R&B',]
```




#### Step 2 - Compare lyrics length

random select 1000 songs and visulize their length comparision

Hip Hop is longer than R&B.
Pop tends to be longer than country songs. 

```{r}
library(dplyr)
library(ngram)
total_four_songs = bind_rows(pop,country,hiphop,RB)

library(stringr)
total_four_songs$word_count <- str_count(total_four_songs$lyrics, '\\s+')+1

```



```{r}
boxplot(word_count~genre,data=total_four_songs, main = "Comparision of Length",col = "orange",
   xlab="genre", ylab="length")
```







#### Step 3 - Compare their high-frequency words - use it to define its topics

Pop and country are simiar. 
If try to make Love, baby .. etc. as new stop words, see what's left .. 

#### 3.1 Pop
```{r}
tidy_pop <- unnest_tokens(pop, output = 'word', token = 'words', input = stemmedwords)%>%
  count(word,sort = TRUE)

# barplot
barplot(tidy_pop[1:10,]$n, las = 2, names.arg = tidy_pop[1:10,]$word,
        col ="pink", main ="Pop - Most frequent words",
        ylab = "Word frequencies")
```

#### 3.2 Country
```{r}
tidy_country <- unnest_tokens(country, output = 'word', token = 'words', input = stemmedwords)%>%
  count(word,sort = TRUE)

# barplot
barplot(tidy_country[1:10,]$n, las = 2, names.arg = tidy_country[1:10,]$word,
        col ="light blue", main ="Country - Most frequent words",
        ylab = "Word frequencies")
```

#### 3.3 HipHops

```{r}
tidy_hiphop <- unnest_tokens(hiphop, output = 'word', token = 'words', input = stemmedwords)%>%
  count(word,sort = TRUE)

# barplot
barplot(tidy_hiphop[1:10,]$n, las = 2, names.arg = tidy_hiphop[1:10,]$word,
        col ="orange", main ="Hiphop - Most frequent words",
        ylab = "Word frequencies")

```


#### 3.4 R&B
```{r}
tidy_RB <- unnest_tokens(RB, output = 'word', token = 'words', input = stemmedwords)%>%
  count(word,sort = TRUE)

# barplot
barplot(tidy_RB[1:10,]$n, las = 2, names.arg = tidy_RB[1:10,]$word,
        col ="black", main ="R&B - Most frequent words",
        ylab = "Word frequencies")
```








### Step 4 - Compare their emotions

Now let's check on the sentiment!

Compare the sentiment score for these 4 different types of songs. 

Hip hop and R&B are more similar, tend to haev stronger emotion variations. Hip-hop tends to have more negative words that R&B. 

Country and Pop are more similar. Their emotions are more gentle, and they both contains balanced emotions. 


```{r}
positive <- get_sentiments("bing") %>%
  filter(sentiment == "positive")

negative <- get_sentiments("bing") %>%
  filter(sentiment == "negative")
```

The top 10 positive words in Pop songs include: 
```{r}
tidy_pop %>%
  semi_join(positive) 
```
The top 10 positive words in Pop songs include: 
```{r}
library(dplyr)
tidy_pop_negative <- tidy_pop%>%
  semi_join(negative) 
head(tidy_pop_negative,10)
```

we count up how many positive and negative words there are in each genre. Let’s find a sentiment score for each word  -- compare setiment along different genres

Referece:https://www.tidytextmining.com/sentiment.html from *Text mining with R*

# sentiment for each genre 


# Combine the first 400 songs of each genre

```{r}
pop_400 = pop[1:400,]
country_400 = country[1:400,]
hiphop_400 = hiphop[1:400,]
RB_400 = RB[1:400,]

total_400 = bind_rows(pop_400,country_400,hiphop_400,RB_400)

tidy_total_400 <- unnest_tokens(total_400, output = 'word', token = 'words', input = stemmedwords)

```


From the diagram, we can see, Hip-hop and R&B tend to have stronger emotions, while country and pop are more gentle.

```{r}
library(tidyr)
bing <- get_sentiments("bing")

songs_sentiment_try <- tidy_total_400 %>%
  inner_join(bing) %>%
  count(genre,song,sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

ggplot(data = songs_sentiment_try, aes(x=song,y=sentiment, fill=genre,color=genre))+
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~genre, ncol = 2, scales = "free_x")

```


# Summary 




### References
Quora (2019) https://www.quora.com/How-do-you-differentiate-between-a-country-song-and-a-pop-song-just-by-listening-to-it


























